Analyzing Strawberry Pricing Strategies Across Canadian Vendors

### 2024-12-02 paper/paper.qmd
add sketches

---
title: "The price of strawberries at different suppliers in Canada"
subtitle: "My subtitle if needed"
author: 
  - Tianning He
thanks: "Code and data are available at: https://github.com/HeTianning/The-price-of-strawberries"
date: 25 November 2024
date-format: long
abstract: "This study investigates the pricing of strawberries across eight major suppliers in Canada: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods. Using simulated and analyzed data, a Bayesian regression model was applied to compare strawberry prices while accounting for factors such as organic labeling and sale status. Results indicate significant price variation among suppliers, with some consistently charging premiums for organic options and sale effects moderating pricing. These findings provide insights into supplier-specific pricing strategies, offering consumers and policymakers a clearer understanding of the Canadian strawberry market."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Strawberries are one of the most important small berries in Canada and the United States(@macoun1919strawberry), and pricing of strawberries in the Canadian retail market varies widely due to factors such as supplier strategies, promotional practices, and temporal trends. This investigation examines the determinants of strawberry prices, focusing on differences between different suppliers, the role of discounts, and seasonal effects. By utilizing a Bayesian linear regression model, the study examines the relationship between key predictors (supplier, original price, promotion, and time) and the current price of strawberries. The Bayesian approach not only provides parameter estimates but also quantifies uncertainty, thus providing a nuanced understanding of price dynamics in competitive and seasonal markets.
The main objective of this analysis is to estimate and explain the factors that influence retail strawberry prices in Canada. The dependent variable current price reflects the Canadian dollar price per pound of strawberries. The key predictors include supplier identity (categorical), which captures supplier-specific pricing strategies; old price (continuous), which represents the original price before promotion, and month (continuous), which captures temporal or seasonal trends(@hodgdon2024current). By modeling these relationships, the study aims to quantify how these variables affect pricing and explore their combined impact on strawberry price changes.
The results show significant differences in supplier-specific pricing strategies, with Metro offering higher prices than other suppliers, consistent with its premium market positioning. Time trends indicate a gradual but consistent increase in strawberry prices over time, possibly reflecting seasonal demand or inflationary pressures. Furthermore, the analysis shows a strong relationship between original and discounted prices, suggesting that promotional pricing generally preserves relative price hierarchies among suppliers. Bayesian models confirm strong convergence and robust posterior estimates, ensuring the reliability of these findings.
Understanding the factors driving strawberry prices is critical for both consumers and retailers. For suppliers, insights into pricing dynamics can inform strategies to optimize revenue while remaining competitive. For consumers, this analysis provides valuable information to navigate the market, highlighting opportunities to save costs by selecting suppliers or purchasing during promotions. More broadly, these findings contribute to the study of the fresh produce market, where pricing is influenced by a complex interaction between supplier strategies, seasonal supply, and consumer behavior. This study lays the foundation for future research that could incorporate additional factors, such as geographic differences or organic certification, to further enrich our understanding of pricing dynamics in Canadian retail.

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
To simulate, test, download, clean, and model data throughout this analysis, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `arrow` [@arrow], `here` [@here], `testthat` [@testthat], and `modelsummary` [@modelsummary]. 
## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

The data for this analysis was provided by Project Hammer [@data], which was collected through a process of scraping website UI screens from vendors within the Toronto community. It contains data on product details (e.g., product name, brand, current and old price, and unit of product) for 8 different grocery stores (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods). Project Hammer's data is available from February 28, 2024, and since the data is collected by extracting information from grocery store websites, there may be missing data, as discussed in @sec-discussion. Through the method of screen scraping, the raw data is intended to measure and reflect grocery store vendor decisions and product pricing trends.
For this analysis, the eight vendors - Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods - and their pricing for strawberries were considered. The variable “price” for each supplier in this analysis reflects the current price of their product at the time the data was downloaded.

@@

An additional variable, old_price, is another continuous measure representing the original price of strawberries before discounts or promotional adjustments. The difference between old_price and current_price can be calculated to assess the magnitude of price reductions, providing insights into sales promotions, markdown strategies, or vendor-specific pricing policies. Both current_price and old_price are essential for understanding vendor behavior and the dynamics of pricing over time.

@@

The primary outcome variable in this study is current_price, which represents the current price of strawberries in Canadian dollars. This continuous variable serves as the dependent variable and captures the prices consumers pay at the point of purchase, reflecting vendor-specific pricing strategies and promotional adjustments. Additionally, old_price is included as a secondary outcome variable, representing the original price before any discounts. The relationship between these two variables highlights the frequency and magnitude of price reductions, offering insights into vendor markdown strategies.

The analysis focuses on understanding how current_price varies with factors such as vendor and month. The categorical variable vendor identifies the supplier and captures differences in pricing strategies across eight vendors, ranging from premium to budget-focused retailers. These variations help reveal how suppliers target different market segments through pricing. The numeric variable month enables the exploration of seasonal trends and temporal fluctuations, reflecting how supply and demand dynamics influence pricing throughout the year.

By examining both current_price and old_price, the study provides a detailed view of promotional activity and vendor behavior. A significant gap between the two prices indicates aggressive promotional strategies, while minimal differences suggest stable pricing policies. Additionally, incorporating the month variable allows for insights into seasonal peaks or dips in strawberry prices, such as higher prices during off-seasons or lower prices during harvest months. Together, these variables offer a comprehensive understanding of strawberry pricing dynamics in Canada’s retail market, helping inform consumer choices and vendor pricing strategies.

@@

2. Original and Current Prices (old_price and current_price)
The dataset includes two continuous price variables: old_price (original price) and current_price (price after any promotions or discounts). The difference between these two variables provides insights into promotional behavior and markdown strategies. Vendors like No Frills and Save-On-Foods typically show small gaps between old_price and current_price, suggesting limited use of deep discounts, whereas Metro and Walmart exhibit larger gaps, reflecting more aggressive promotional activity. Boxplots comparing old_price and current_price by vendor demonstrate how discounts influence consumer pricing perceptions and vendor competition.

@@

4. Combined Factors: Vendor, Price, and Seasonality
The interplay of vendor, old_price, current_price, and month allows for a more comprehensive analysis of pricing strategies. For example, vendors like Metro may use larger discounts (old_price vs. current_price) during peak seasons to maintain competitiveness, while budget vendors like No Frills may maintain consistent pricing throughout the year. A combined analysis using bar charts and line plots highlights how vendor strategies vary with time, providing insights into the dynamic nature of strawberry pricing.

The inclusion of variables like vendor, old_price, current_price, and month enriches the analysis by offering multiple perspectives on pricing strategies. Vendor captures the segmentation of suppliers into premium and budget categories, reflecting strategic decisions about market positioning. The dual price variables (old_price and current_price) allow for a deeper understanding of promotional activity and highlight the extent to which vendors rely on discounts to influence consumer behavior.

@@

```{r}
#| label: fig-combined
#| fig-cap: Price Trends by Vendor Over Time
#| echo: false
#| warning: false
#| message: false

# Model

@@

The chosen model is a **Bayesian linear regression model**, which builds on the simplicity of traditional linear regression while incorporating the Bayesian framework. This allows for probabilistic insights and robust uncertainty quantification, making it well-suited for analyzing the factors driving strawberry prices.

@@

#### Why Use a Bayesian Linear Regression Model?
A Bayesian linear regression model is a suitable approach for this analysis because it balances simplicity, interpretability, and flexibility while quantifying uncertainty in parameter estimates. This model is particularly effective for exploring the continuous outcome variable, current_price, and its relationship with predictors such as vendor, month, and original price (old_price). Unlike traditional linear regression, which assumes a deterministic relationship between predictors and the outcome, the Bayesian framework provides posterior distributions for all parameters, enabling richer and more nuanced inferences. This approach allows us to incorporate prior knowledge, account for uncertainty, and provide probabilistic insights into strawberry pricing.

The Bayesian framework also handles sparse or imbalanced data effectively by leveraging prior distributions to stabilize parameter estimates. For example, when certain vendors have fewer observations, the priors prevent overfitting and provide regularized estimates, making the model robust to variability in data distribution. This ensures reliable insights into vendor-specific pricing strategies, seasonal trends, and the influence of baseline (old_price) on observed (current_price) prices.
#### Benefits of Using Bayesian Regression

@@

# Results

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

@@

# Discussion

@@

## The Role of Original Price in Promotional Pricing
The strong and significant relationship between old_price and current_price highlights the structured nature of promotional pricing strategies. Higher original prices are associated with proportionally higher current prices, even after discounts or sales are applied. This suggests that vendors maintain pricing hierarchies across their product offerings, preserving the perceived value of premium products even during promotional periods. For vendors, this finding underscores the importance of setting initial prices strategically, as they serve as an anchor for both regular and discounted pricing. For consumers, it suggests that discounts may not always reflect deep value but instead preserve the relative pricing order among vendors or product lines. This structured approach to promotional pricing also indicates that vendors likely use data-driven algorithms or market insights to optimize discounts while maintaining brand positioning. Future research could explore the behavioral aspects of how consumers respond to these pricing strategies, shedding light on how perceptions of value and fairness influence purchasing decisions during sales events.

@@

# Additional data details

# Model details {#sec-model-details}

>>>>

---
title: "Analyzing Strawberry Pricing Strategies Across Canadian Vendors"
subtitle: "A Data-Driven Investigation of Seasonal Trends, Vendor Variability, and Promotional Impacts on Grocery Pricing"
author: 
  - Tianning He
thanks: "Code and data are available at: https://github.com/HeTianning/The-price-of-strawberries"
date: 25 November 2024
date-format: long
abstract: "This study investigates the pricing of strawberries across eight major suppliers in Canada: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods. Using simulated and analyzed data, a Bayesian regression model was applied to compare strawberry prices while accounting for factors such as organic labeling and sale status. Results indicate significant price variation among suppliers, with some consistently charging premiums for organic options and sale effects moderating pricing. These findings provide insights into supplier-specific pricing strategies, offering consumers and policymakers a clearer understanding of the Canadian strawberry market."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Strawberries are one of the most important small berries in Canada and the United States(@macoun1919strawberry), and pricing of strawberries in the Canadian retail market varies widely due to factors such as supplier strategies, promotional practices, and temporal trends. This investigation examines the determinants of strawberry prices, focusing on differences between different suppliers, the role of discounts, and seasonal effects. By utilizing a Bayesian linear regression model, the study examines the relationship between key predictors (supplier, original price, promotion, and time) and the current price of strawberries. The Bayesian approach not only provides parameter estimates but also quantifies uncertainty, thus providing a nuanced understanding of price dynamics in competitive and seasonal markets.

The main objective of this analysis is to estimate and explain the factors that influence retail strawberry prices in Canada. The dependent variable current price reflects the Canadian dollar price per pound of strawberries. The key predictors include supplier identity (categorical), which captures supplier-specific pricing strategies; old price (continuous), which represents the original price before promotion, and month (continuous), which captures temporal or seasonal trends(@hodgdon2024current). By modeling these relationships, the study aims to quantify how these variables affect pricing and explore their combined impact on strawberry price changes.

The results show significant differences in supplier-specific pricing strategies, with Metro offering higher prices than other suppliers, consistent with its premium market positioning. Time trends indicate a gradual but consistent increase in strawberry prices over time, possibly reflecting seasonal demand or inflationary pressures. Furthermore, the analysis shows a strong relationship between original and discounted prices, suggesting that promotional pricing generally preserves relative price hierarchies among suppliers. Bayesian models confirm strong convergence and robust posterior estimates, ensuring the reliability of these findings.

Understanding the factors driving strawberry prices is critical for both consumers and retailers. For suppliers, insights into pricing dynamics can inform strategies to optimize revenue while remaining competitive. For consumers, this analysis provides valuable information to navigate the market, highlighting opportunities to save costs by selecting suppliers or purchasing during promotions. More broadly, these findings contribute to the study of the fresh produce market, where pricing is influenced by a complex interaction between supplier strategies, seasonal supply, and consumer behavior. This study lays the foundation for future research that could incorporate additional factors, such as geographic differences or organic certification, to further enrich our understanding of pricing dynamics in Canadian retail.

The rest of the paper is structured as follows. @sec-data describes the data collection process, including data sources, sampling strategy, and preprocessing steps taken to ensure data quality. @sec-model outlines the methodological framework, focusing on the Bayesian linear regression model, its formulation, and diagnostic checks. @sec-result presents the main findings of the analysis, focusing on specific supplier pricing strategies, seasonal trends, and the impact of sales and promotions on strawberry prices. @sec-discussion discusses the implications of these findings, placing them in the broader context of consumer behavior and supplier strategies in the fresh produce market.

# Data {#sec-data}
To simulate, test, download, clean, and model data throughout this analysis, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `arrow` [@arrow], `here` [@here], `testthat` [@testthat], and `modelsummary` [@modelsummary]. 
## Overview
The data for this analysis was provided by Project Hammer [@filipp2024hammer], which was collected through a process of scraping website UI screens from vendors within the Toronto community. It contains data on product details (e.g., product name, brand, current and old price, and unit of product) for 8 different grocery stores (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods). Project Hammer's data is available from February 28, 2024, and since the data is collected by extracting information from grocery store websites, there may be missing data, as discussed in appendix. Through the method of screen scraping, the raw data is intended to measure and reflect grocery store vendor decisions and product pricing trends.
For this analysis, the eight vendors - Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods - and their pricing for strawberries were considered. The variable “price” for each supplier in this analysis reflects the current price of their product at the time the data was downloaded.

@@

An additional variable, old price, is another continuous measure representing the original price of strawberries before discounts or promotional adjustments. The difference between old price and current price can be calculated to assess the magnitude of price reductions, providing insights into sales promotions, markdown strategies, or vendor-specific pricing policies. Both current price and old price are essential for understanding vendor behavior and the dynamics of pricing over time.

@@

The primary outcome variable in this study is current price, which represents the current price of strawberries in Canadian dollars. This continuous variable serves as the dependent variable and captures the prices consumers pay at the point of purchase, reflecting vendor-specific pricing strategies and promotional adjustments. Additionally, old price is included as a secondary outcome variable, representing the original price before any discounts. The relationship between these two variables highlights the frequency and magnitude of price reductions, offering insights into vendor markdown strategies.

The analysis focuses on understanding how current price varies with factors such as vendor and month. The categorical variable vendor identifies the supplier and captures differences in pricing strategies across eight vendors, ranging from premium to budget-focused retailers. These variations help reveal how suppliers target different market segments through pricing. The numeric variable month enables the exploration of seasonal trends and temporal fluctuations, reflecting how supply and demand dynamics influence pricing throughout the year.

By examining both current price and old price, the study provides a detailed view of promotional activity and vendor behavior. A significant gap between the two prices indicates aggressive promotional strategies, while minimal differences suggest stable pricing policies. Additionally, incorporating the month variable allows for insights into seasonal peaks or dips in strawberry prices, such as higher prices during off-seasons or lower prices during harvest months. Together, these variables offer a comprehensive understanding of strawberry pricing dynamics in Canada’s retail market, helping inform consumer choices and vendor pricing strategies.

@@

2. Old and Current Prices
The dataset includes two continuous price variables: old price (original price) and current price (price after any promotions or discounts). The difference between these two variables provides insights into promotional behavior and markdown strategies. Vendors like No Frills and Save-On-Foods typically show small gaps between old price and current price, suggesting limited use of deep discounts, whereas Metro and Walmart exhibit larger gaps, reflecting more aggressive promotional activity. Boxplots comparing old price and current price by vendor demonstrate how discounts influence consumer pricing perceptions and vendor competition.

@@

4. Combined Factors: Vendor, Price, and Seasonality
The interplay of vendor, old price, current price, and month allows for a more comprehensive analysis of pricing strategies. For example, vendors like Metro may use larger discounts (old price vs. current price) during peak seasons to maintain competitiveness, while budget vendors like No Frills may maintain consistent pricing throughout the year. A combined analysis using bar charts and line plots highlights how vendor strategies vary with time, providing insights into the dynamic nature of strawberry pricing.

The inclusion of variables like vendor, old price, current price, and month enriches the analysis by offering multiple perspectives on pricing strategies. Vendor captures the segmentation of suppliers into premium and budget categories, reflecting strategic decisions about market positioning. The dual price variables (old price and current price) allow for a deeper understanding of promotional activity and highlight the extent to which vendors rely on discounts to influence consumer behavior.

@@

# Model{#sec-model}

@@

The chosen model is a **Bayesian linear regression model**, which builds on the simplicity of traditional linear regression while incorporating the Bayesian framework. This allows for probabilistic insights and robust uncertainty quantification, making it well-suited for analyzing the factors driving strawberry prices. Further background details and diagnostics are included in [Appendix -@sec-model-details].

@@

#### Why Use a Bayesian Linear Regression Model?
A Bayesian linear regression model is a suitable approach for this analysis because it balances simplicity, interpretability, and flexibility while quantifying uncertainty in parameter estimates. This model is particularly effective for exploring the continuous outcome variable, current_price, and its relationship with predictors such as vendor, month, and original price (old price). Unlike traditional linear regression, which assumes a deterministic relationship between predictors and the outcome, the Bayesian framework provides posterior distributions for all parameters, enabling richer and more nuanced inferences. This approach allows us to incorporate prior knowledge, account for uncertainty, and provide probabilistic insights into strawberry pricing.

The Bayesian framework also handles sparse or imbalanced data effectively by leveraging prior distributions to stabilize parameter estimates. For example, when certain vendors have fewer observations, the priors prevent overfitting and provide regularized estimates, making the model robust to variability in data distribution. This ensures reliable insights into vendor-specific pricing strategies, seasonal trends, and the influence of baseline (old price) on observed (current price) prices.
#### Benefits of Using Bayesian Regression

@@

# Results{#sec-result}

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory model based on strawberry prices"
#| warning: false

@@

# Discussion{#sec-discussion}

@@

## The Role of Original Price in Promotional Pricing
The strong and significant relationship between old price and current price highlights the structured nature of promotional pricing strategies. Higher original prices are associated with proportionally higher current prices, even after discounts or sales are applied. This suggests that vendors maintain pricing hierarchies across their product offerings, preserving the perceived value of premium products even during promotional periods. For vendors, this finding underscores the importance of setting initial prices strategically, as they serve as an anchor for both regular and discounted pricing. For consumers, it suggests that discounts may not always reflect deep value but instead preserve the relative pricing order among vendors or product lines. This structured approach to promotional pricing also indicates that vendors likely use data-driven algorithms or market insights to optimize discounts while maintaining brand positioning. Future research could explore the behavioral aspects of how consumers respond to these pricing strategies, shedding light on how perceptions of value and fairness influence purchasing decisions during sales events.

@@

## Additional data details
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary Statistics and Distribution Overview of the Analysis Dataset"
#| warning: false

datasummary_skim(analysis_dataset)

```
###Data Source and Cleaning Process
The data for this study comes from the Project Hammer website, an open data platform centered on grocery prices from multiple suppliers in Canada. The platform captures multidimensional data including product prices, supplier information, promotional discounts, etc. through web scraping technology, and combines manual verification to ensure its reliability. These data provide price dynamics in different regions and suppliers in Canada, laying the foundation for studying the fluctuations and market strategies of strawberry prices.

In order to meet the analysis needs, the data was cleaned in multiple steps. First, the rows with missing values in the current price and old price variables were deleted because these prices are the core of the study. If the price records are incomplete, the analysis results will be distorted. In addition, for the vendor and month variables, it was confirmed that there were no missing values to ensure the integrity of the supplier and time information.

Secondly, the interquartile range (IQR) method was used to detect and process outliers. The focus of the cleaning was on identifying and eliminating possible entry errors (such as extremely low or high prices). However, legitimate outliers (such as high prices for high-end strawberries) were retained to ensure the authenticity of the data. At the same time, all price fields were standardized and prices were recorded uniformly as the price per pound of strawberries to facilitate horizontal comparison between suppliers.

###The strengths and limitations of data
On the plus side, the data is diverse, covering both national and regional suppliers, including Voila, Metro, and Walmart, among others. This provides a unique perspective on the differences between different market positioning and pricing strategies. In addition, the data cover observations across multiple months, providing a solid basis for exploring seasonal fluctuations in strawberry prices. Simultaneous recording of the price variables current price and old price allows studying the impact of discount and promotion strategies on the final price.

Although the data has significant advantages, there are also some limitations. First, the data focuses primarily on urban and suburban providers, and pricing trends in rural areas may be understated. Second, web scraping technology captures static snapshots and is difficult to record real-time price fluctuations, such as flash sales or limited-time discounts. In addition, some records lack explicit promotion status markers (such as whether the product is discounted), which limits detailed analysis of discount strategies. Finally, the density of records across months may be uneven across different providers, which may introduce bias in the temporal analysis.

##Data Analysis Outlook
Based on the cleaned data, the analysis can focus on several key directions. First, by combining the current price and old price variables, we can delve into the impact of promotions on pricing strategies. For example, we can quantify the magnitude of price fluctuations during promotions for different suppliers and their potential impact on consumer behavior. Second, through the month variable, we can capture the seasonal fluctuations in strawberry prices. This will help us understand how supply and demand adjust with seasonal changes and provide an important basis for predicting future price trends.

In addition, using the vendor variable, we can compare the pricing strategies of different suppliers. The study may reveal whether high-end suppliers (such as Metro) are positioned in the high-end market through higher prices, while low-price suppliers (such as Walmart and Save-On-Foods) attract more price-sensitive consumers through discounts. These findings will provide in-depth insights into understanding the dynamics of the Canadian strawberry market, while providing data support for policymakers and retailers to optimize market strategies.

Overall, the cleaned dataset provides a valuable resource for in-depth study of strawberry price fluctuations and supplier strategies. Despite limitations such as geographic coverage bias and insufficient capture of dynamic prices, the diversity and time span of the data ensure the wide applicability of the analysis. By combining the analysis of promotions, seasonal fluctuations, and supplier differences, the study reveals complex pricing patterns and competitive strategies in the Canadian market.

## Further Background Details and Diagnostics of the Model {#sec-model-details}
###Model Background and Justification
The Bayesian linear regression model provides a probabilistic framework to understand the determinants of strawberry pricing across Canadian vendors. This model was chosen for its ability to incorporate prior knowledge, handle parameter uncertainty, and deliver intuitive, probabilistic interpretations of results. Unlike traditional frequentist regression, Bayesian regression generates posterior distributions for all parameters, offering insights into the uncertainty and variability of each predictor's effect.

The primary dependent variable in the model is `current_price`, representing the observed price per pound of strawberries. Key predictors include vendor identity (`vendor`), original price (`old_price`), sale status (`sale_status`), and temporal trends (`month`). These variables capture the diverse factors influencing pricing, such as competitive strategies among vendors, promotional effects, and seasonal supply-demand dynamics.

The model specification includes normal priors for regression coefficients (\(\beta\)) and an exponential prior for the residual standard deviation (\(\sigma\)). This setup reflects reasonable assumptions about the effects of predictors, such as small, centered effects with moderate variability. For instance, the prior for \(\beta_{\text{vendor}}\) assumes no extreme pricing differences between vendors, while allowing the data to dominate posterior updates.

### Model Diagnostics and Validation
#### Posterior Predictive Checks
Posterior predictive checks (pp_checks) compare the observed distribution of `current_price` with simulated data generated from the model. The goal is to assess whether the model adequately captures the variability and central tendencies of the observed data. For this analysis, posterior predictive plots revealed close alignment between observed and predicted prices, indicating a good fit. Minor discrepancies at the tails suggest that the model may not fully capture extreme price outliers, which could be addressed by adding robust regression components or alternative distributions.

#### Trace Plots
Trace plots visualize the behavior of the Markov Chain Monte Carlo (MCMC) sampling process for each parameter. Convergence is achieved when chains are stationary and overlap across iterations. For this model, trace plots showed consistent mixing and no evidence of divergence, confirming that the chains have converged and are exploring the posterior distributions effectively.

#### Rhat Values
The Rhat statistic evaluates chain convergence across parallel MCMC simulations. All Rhat values for the parameters in this model were close to 1.0, indicating excellent convergence. This result provides confidence that the posterior distributions are reliable and free from sampling biases.

#### Residual Analysis
Residual plots were analyzed to assess the assumptions of normality and independence of errors. The residuals appeared approximately normal, with minor deviations for extreme values. Temporal residual plots showed no significant autocorrelation, suggesting that the inclusion of the `month` variable sufficiently accounted for seasonal trends.

#### Model Fit Metrics
- **\( R^2 \)** and **Adjusted \( R^2 \)**: The model explained approximately 13.9% of the variance in strawberry prices, reflecting the complexity of pricing dynamics influenced by unobserved factors like geographic location and consumer preferences.
- **Root Mean Square Error (RMSE)**: An RMSE of 11.79 indicates moderate predictive accuracy, which aligns with expectations given the variability in pricing strategies across vendors.
- **Leave-One-Out Cross-Validation (LOO-CV)**: The LOO-CV score confirmed the model's predictive consistency and supported its suitability for analyzing strawberry pricing.

## Areas for Model Improvement

### Interaction Terms
Interaction terms, such as between `vendor` and `sale_status`, could capture more nuanced effects. For example, some vendors might offer larger discounts during promotions than others, which is not explicitly modeled here.

### Non-Linearity
The assumption of linear relationships between predictors and `current_price` may oversimplify the effects of variables like `month`, which could exhibit non-linear seasonal trends. Incorporating generalized additive models (GAMs) or polynomial terms for time could enhance the model’s flexibility.

### Alternative Priors
While normal priors for coefficients are intuitive and effective, sensitivity analysis could explore alternative priors, such as student-t distributions, which are more robust to outliers.

### Additional Predictors
Including variables such as geographic location (`region`) or packaging type (`packaging`) could improve the explanatory power of the model and account for unobserved heterogeneity.

## Summary

The Bayesian linear regression model provides a robust framework for understanding strawberry pricing dynamics. Diagnostic checks, including posterior predictive plots, trace plots, and Rhat values, confirm the model's validity and reliability. While the model captures the major drivers of price variability, areas for improvement include modeling interactions, addressing non-linearity, and incorporating additional predictors. These enhancements could refine the analysis and provide deeper insights into vendor strategies and market trends.
